---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# eye

This package will help you work with eye data. 

# Features

## Eye stuff

### Conversion of visual acuity notations 

### [Easy count of patients and eyes](#count-patients-and-eyes)

## Beyond eye stuff

### Stats functions
- Show common summary statistics
- Calculate 2d probability contours for the use in ggplot2

### Convenience functions: 
- Save your data frame as csv 
- Calculate age
- Anonymize your data frame

### ggplot2 extensions 
- geom_trail: Make a "base plot type = b" with ggplot

# Examples
## Eye stuff
### Count patients and eyes
```{r example1, warning=FALSE, message=FALSE, results='hide'}

```
### Get basic statistics
```{r mpstats, warning=FALSE, message=FALSE}

```
### Plot probability contours
```{r example2, warning=FALSE, message=FALSE, results='hide'}

```
### Make a "base plot type = b" with ggplot
```{r trail}

```

### Calculate age
```{r age, warning=FALSE, message=FALSE}

```
### Anonymize your data 
```{r anonymize, warning=FALSE, message=FALSE}

```
### Export your data to csv 
```{r csv, warning=FALSE, message=FALSE}

```
## Installation 
```{r, eval=FALSE}
# for the development version 
devtools::install_github(tjebo/eye)
```

## Sources
Containing data from Fasler et al.. [data](https://bmjopen.bmj.com/content/9/6/e027441)

