---
output:
  md_document:
    variant: gfm
bibliography: eye.bib
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(eye)
```

# eye

See your data with a new eye. 

# Features
## Only about the eye
- [Easy count of patients and eyes](#count-patients-and-eyes)
- [Conversion of visual acuity notations](#va-conversion)

## Beyond the eye
### Stats functions
- [Show common summary statistics](#basic-statistics)
- [Calculate 2d probability contours for the use in ggplot2](#probability-contours)

### Convenience functions: 
- [Easy save of your data frame as csv](#save-to-csv)
- [Calculate age](#calculate-age)

### ggplot2 extensions 
- [geom_trail: A base plot type = "b" equivalent for ggplot2](#geom_trail)

## AMD data 
Real life anonymised human subject data from a large cohort of patients with treatment-naive neovascular age-related macular degeneration (AMD) that received intravitreal anti-VEGF therapy. The data is from Moorfields Eye Hospital, London, UK and was made available [on a public repository](https://datadryad.org/stash/dataset/doi:10.5061/dryad.97r9289). 

To reference this data in your publication, please kindly cite the corresponding article by Fasler and colleagues.[@fasler] 

# Get eye 
```{r, eval=FALSE}
# for the development version 
devtools::install_github("tjebo/eye")
```

# Examples
## Eye stuff
### Count patients and eyes
```{r eyes, warning=TRUE, message=FALSE}
eyes(amd)
```

### VA conversion
```{r va, warning=FALSE, message=FALSE, results='hide'}

```

## Beyond eye stuff
### Common statistics
```{r stats, warning=FALSE, message=FALSE}
see(amd$BaselineAge, dec = 1)

see(amd[c("BaselineAge", "VA_ETDRS_Letters", "FollowupDays")])
```

### Probability contours
```{r prob, warning=FALSE, message=FALSE, results='hide'}
library(ggplot2)

set.seed(1)
n=100
foo <- data.frame(x=rnorm(n, 0, 1), y=rnorm(n, 0, 1))

df_contours <- dplyr::bind_rows(
  purrr::map(seq(0.2, 0.8, 0.2), function(p) prob_contour(foo, prob = p))
)

ggplot() +
  geom_point(data = foo, aes(x = x, y = y)) +
  geom_polygon(data = df_contours, aes(x = x, y = y, color = prob), fill = NA) +
  scale_color_brewer(name = "Probs", palette = "Set1")
```

### Calculate age
Age calculated in years. Periods or duration possible. If only one argument, the age today.
```{r age, warning=FALSE, message=FALSE}
age("1984-10-16")

dob <-  c("1984-10-16", "2000-01-01")
test_date <-  as.Date(dob) + c(15000, 20000)

age(dob, test_date)
```

### Save to csv 
This is a convenience wrapper around write.csv
```{r csv, warning=FALSE, message=FALSE, eval = FALSE}
csv(amd)
```

## ggplot2 extensions
### geom_trail
A base plot type = "b" equivalent for ggplot. Works also with text!
```{r trail}
library(ggplot2)

ggplot(pressure, aes(temperature, pressure)) +
  geom_ribbon(aes(ymin = pressure - 50, ymax = pressure + 50), alpha = 0.2) +
  geom_trail()
```

# References


