---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# eye

This package will help you work with eye data. 

# Features
## Only about the eye
### [Easy count of patients and eyes](#count-patients-and-eyes)
### [Conversion of visual acuity notations](#va-conversion)

## Beyond the eye
### Stats functions
- [Show common summary statistics](#basic-statistics)
- [Calculate 2d probability contours for the use in ggplot2](#probability-contours)
### Convenience functions: 
- Save your data frame as csv 
- Calculate age
- Anonymize your data frame
### ggplot2 extensions 
- [geom_trail: A base plot type = "b" equivalent for ggplot2](#geom_trail)

# Examples
## Eye stuff
### Count patients and eyes
```{r eyes, warning=FALSE, message=FALSE, results='hide'}

```
### VA conversion
```{r va, warning=FALSE, message=FALSE, results='hide'}

```
## Beyond eye stuff
### Basic statistics
```{r stats, warning=FALSE, message=FALSE}

```
### Probability contours
```{r prob, warning=FALSE, message=FALSE, results='hide'}

```
## ggplot2 extensions
### geom_trail
A base plot type = "b" equivalent for ggplot. Works also with text!
```{r trail}

```

### Calculate age
```{r age, warning=FALSE, message=FALSE}

```
### Anonymize your data 
```{r anonymize, warning=FALSE, message=FALSE}

```
### Export your data to csv 
```{r csv, warning=FALSE, message=FALSE}

```
## Installation 
```{r, eval=FALSE}
# for the development version 
devtools::install_github(tjebo/eye)
```

## Sources
Containing data from Fasler et al.. [data](https://bmjopen.bmj.com/content/9/6/e027441)

